---
title: "Assignment 4 Ellis ML 64060"
author: "Samantha Ellis"
date: "2023-10-26"
output: html_document
---
Question A.
```{r}
##Looking only at nummeric variables
NumericPharm<- Pharmaceuticals %>% select(3:11)
##Scaling the dataset
ScaledNumericPharm<- scale(NumericPharm)
##Determining the number of clusters needed
ClustScalNumPharm<- (nrow(ScaledNumericPharm)-1)*sum(apply(ScaledNumericPharm,2,var))
ClustScalNumPharm
for (i in 2:15)ClustScalNumPharm[i]<- sum(kmeans(ScaledNumericPharm, centers = i)$withinss)
ClustScalNumPharm
plot(1:15, ClustScalNumPharm, type = "b", xlab = "Number of Clusters", ylab = "Within groups sum of squares")
##The visual shows a drop around 5, so that would be appropriate amount of clusters for this dataset
```
I initially attempted using functions using the flexclust package and the factoextra package, but those are no longer available within RStudio, so I had to find alternative functions and methods.
I used the Elbow Method to determine number of clusters.




Question B.
```{r}
##The mean of each cluster by variable
KMEANClustPharm<- kmeans(ScaledNumericPharm, 5)
aggregate(ScaledNumericPharm, by=list(KMEANClustPharm$cluster),FUN=mean)
```
Cluster 1 is the lowest for Market capitalization, highest for Beta, lowest for Price/earnings ratio, lowest for asset turnover, and highest for leverage.
Cluster 2 has the highest return of equity and highest return of assets, but the lowest leverage. 
Cluster 3 has the highest Market capitalization, asset turnover, estimated revenue growth, and net profit margin. 
Cluster 4 the highest price/earnings ratio, but the lowest return of assets, return of equity, and net profit margin. 
Cluster 5 has the lowest Beta and lowest estimated revenue growth. 

This follows the silhouette method. Again, since some packages from the lecture are outdated, I found other functions to display the clusters with similar results.




Question C. 
```{r}
clusplot(ScaledNumericPharm, KMEANClustPharm$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```
Cluster 3 is the highest in four of the nine variables, and is not the lowest for any of the other variables. Cluster 5 is the lowest for two of the variables and is not the highest for any of them. 
Visually, clusters 2, 4, and 5, have similar shapes and spreads (ovals that skew right and downwards), while Clusters 1 and 3 have unique spreads. 
Visually, Cluster 3, is the farthest outlier, and only contains two of the 21 pharmaceutical companies. 










Question D.
Cluster 1 should be labelled as Highest Beta, Leverage, and Lowest Market Cap, PER, AT.
Cluster 2 should be labelled as Highest ROE, ROA, and Lowest leverage.
Cluster 3 should be labelled as Highest Market Cap, AT, ERG, NPM.
Cluster 4 should be labelled as Highest PER, and Lowest ROE, ROA, NPM.
Cluster 5 should be labelled as Lowest Beta, ERG.

